---
title: "econ_4"
author: "Lydia Bleifuss"
date: "5/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#set up 
library(tidyverse)
library(rootSolve)
library(janitor)
```

```{r include= FALSE}
#read in data
abate_data <- read.csv("HW4Data.csv")
abate_data
```

### Part 1. Functional form for the marginal cost of abatement for each sector
Write down a plausible functional form for the marginal cost of abatement for sector A. Use regression analysis to estimate the parameters of that function. Repeating this for sectors B, C, and D will give you a model of the marginal cost of abatement function for each sector. How well do your models fit the data for each sector? You may need to experiment with different functional forms. Produce a plot of the estimated marginal abatement cost functions in all four sectors (this plot should go in your memo).

# Sector A
```{r}

#Explore the data for Sector A

plot_sector_a <- ggplot(abate_data, aes(x = Abatement, y = MC_A ))+
  geom_point()

plot_sector_a # The data for Sector A appears linear

# Use linear regression to find the function for Sector A

sector_a_lm <- lm(MC_A ~ Abatement, data = abate_data)

sector_a_lm

# Plot the linear model to see if it fits the data well 

sector_a_lm_plot <- ggplot(abate_data, aes(x=Abatement, y=MC_A)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = y~x)

sector_a_lm_plot # The model seems to fit the data

# Save the regression coefficients

sector_a_intercept <- sector_a_lm$coefficients[1]
sector_a_slope <- sector_a_lm$coefficients[2]


# Write the marginal abatement cost function for Sector A

sector_a_function <- function(x){sector_a_intercept + sector_a_slope*x}

```

$Sector~A~Marginal~Abatement~Cost = `r sector_a_intercept`+`r sector_a_slope`*Q$ 

# Sector B 

```{r}

#Explore the data for Sector B

plot_sector_b <- ggplot(abate_data, aes(x = Abatement, y = MC_B ))+
  geom_point()

plot_sector_b # The data for sector B appears linear

# Use linear regression to find the function for Sector B

sector_b_lm <- lm(MC_B ~ Abatement, data = abate_data)

sector_b_lm

# Plot the linear model to see if it fits the data well 

sector_b_lm_plot <- ggplot(abate_data, aes(x=Abatement, y=MC_B)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = y~x)

sector_b_lm_plot # The model seems to fit the data

# Save the regression coefficients

sector_b_intercept <- sector_b_lm$coefficients[1]
sector_b_slope <- sector_b_lm$coefficients[2]

# Write the marginal abatement cost function for Sector B

sector_b_function <- function(x){sector_b_intercept + sector_b_slope*x}

```

$Sector~B~Marginal~Abatement~Cost = `r sector_b_intercept`+`r sector_b_slope`*Q$ 

# Sector C

```{r}

#Explore the data for Sector C

plot_sector_c <- ggplot(abate_data, aes(x = Abatement, y = MC_C ))+
  geom_point()

plot_sector_c # The data for sector C may be linear or exponential? Sam is using linear regression

# Use linear regression to find the function for Sector C

sector_c_lm <- lm(MC_C ~ Abatement, data = abate_data)

sector_c_lm

# Plot the linear model to see if it fits the data well 

sector_c_lm_plot <- ggplot(abate_data, aes(x=Abatement, y=MC_C)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = y~x)

sector_c_lm_plot # The model seems to fit the data

# Save the regression coefficients

sector_c_intercept <- sector_c_lm$coefficients[1]
sector_c_slope <- sector_c_lm$coefficients[2]

# Write the marginal abatement cost function for Sector C

sector_c_function <- function(x){sector_c_intercept + sector_c_slope*x}

```

$Sector~C~Marginal~Abatement~Cost = `r sector_c_intercept`+`r sector_c_slope`*Q$ 

# Sector D 

```{r}

#Explore the data for Sector D

plot_sector_d <- ggplot(abate_data, aes(x = Abatement, y = MC_D ))+
  geom_point()

plot_sector_d # The data for sector D appears mostly linear 

# Use linear regression to find the function for Sector S

sector_d_lm <- lm(MC_D ~ Abatement, data = abate_data)

sector_d_lm

# Plot the linear model to see if it fits the data well 

sector_d_lm_plot <- ggplot(abate_data, aes(x=Abatement, y=MC_D)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = y~x)

sector_d_lm_plot # The model seems to fit the data

# Save the regression coefficients

sector_d_intercept <- sector_d_lm$coefficients[1]
sector_d_slope <- sector_d_lm$coefficients[2]

# Write the marginal abatement cost function for Sector D

sector_d_function <- function(x){sector_d_intercept + sector_d_slope*x}

```

$Sector~D~Marginal~Abatement~Cost = `r sector_d_intercept`+`r sector_d_slope`*Q$ 

# Plot the marginal abatment cost functions for all four sectors on one graph

```{r}

all_sectors_plot <- ggplot(data.frame(x = c(0, 200))) +
  stat_function(fun = sector_a_function, aes(color = "A")) +
  stat_function(fun = sector_b_function, aes(color = "B")) +
  stat_function(fun = sector_c_function, aes(color = "C")) +
  stat_function(fun = sector_d_function, aes(color = "D")) +
  scale_y_continuous(limits= c(0,120), expand = c(0,0)) +
  scale_x_continuous(limits= c(0, 200), expand = c(0,0)) + 
  labs(fill = "", color = "Sector") +
  xlab("\n Carbon Abatement (tons)") + 
  ylab("Cost \n($/ton) \n") + 
  ggtitle("Marginal Cost of Carbon Abatement in Four Economic Sectors\n") + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(angle=0, hjust = 0.5, vjust = 0.5), 
        axis.title.x = element_text(angle=0, hjust = 0.5, vjust = 0.5) ) +
  theme_classic() + scale_colour_manual("Sector", values = c("chartreuse4", "goldenrod1","dodgerblue2","red3"))

all_sectors_plot
```

### Part 2. Derive a demand curve for each sector
Using these models and the current level of carbon emissions, derive each sector’s demand curve for carbon emissions. In other words, how much would each sector be willing to pay for the right to pollute the first unit, second unit, etc? Draw these demand curves on a graph. Which sector is willing to pay the most for the first unit of carbon emissions?

```{r}
# To derive the demand curve for each sector we must determine the y intercept (abatement cost of 180th unit is the same as the cost of the first unit of emission)

# use new y intercept to form y = mx + b, or the demand curve for carbon emissions for each sector. Then you are able to find the area under the demand curve, willingness to pay? 

y_int_a <- sector_a_function(180)
y_int_a #95.18706 (WTP for first unit)
y_int_b <- sector_b_function(200)
y_int_b #49.06656 (WTP for first unit)
y_int_c <- sector_c_function(220)
y_int_c #160.7868 (WTP for first unit)
y_int_d <- sector_d_function(300)
y_int_d #87.66577  (WTP for first unit)

# Demand curve eqution for each sector (= intercept + slope*abatement)

#Sector a slope (180)
a_dem_slope <- (0-y_int_a)/(180-0)
a_dem_slope

#Sector a demand curve
a_dem_curve <- function(x){
  y_int_a + a_dem_slope*x
}
a_dem_curve

curve(a_dem_curve, from=0, to=180, xlab = "x", ylab = "y")

#Sector b slope (200)
b_dem_slope <- (0-y_int_b)/(200-0)
b_dem_slope

#Sector b demand curve
b_dem_curve <- function(x){
  y_int_b + b_dem_slope*x
}
b_dem_curve

#Sector c slope (220)
c_dem_slope <- (0-y_int_c)/(220-0)
c_dem_slope

#Sector b demand curve
c_dem_curve <- function(x){
  y_int_c + c_dem_slope*x
}
c_dem_curve

#Sector d slope (300) 
d_dem_slope <- (0-y_int_d)/(300-0)
d_dem_slope

#Sector d demand curve
d_dem_curve <- function(x){
  y_int_d + d_dem_slope*x
}
d_dem_curve

```

```{r}
all_sectors_demand <- ggplot(data.frame(x = c(0, 200))) +
  stat_function(fun = a_dem_curve, aes(color = "A")) +
  stat_function(fun = b_dem_curve, aes(color = "B")) +
  stat_function(fun = c_dem_curve, aes(color = "C")) +
  stat_function(fun = d_dem_curve, aes(color = "D")) +
  scale_y_continuous(limits= c(0,200), expand = c(0,0)) +
  scale_x_continuous(limits= c(0, 350), expand = c(0,0)) + 
  labs(fill = "", color = "Sector") +
  xlab("\n Carbon Emissions (tons)") + 
  ylab("Cost \n($/ton) \n") + 
  ggtitle("Marginal Willingness to Pay in Four Economic Sectors for Carbon Emissions\n") + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(angle=0, hjust = 0.5, vjust = 0.5), 
        axis.title.x = element_text(angle=0, hjust = 0.5, vjust = 0.5) ) +
  theme_classic() + scale_colour_manual("Sector", values = c("chartreuse4", "goldenrod1","dodgerblue2","red3"))

all_sectors_demand

```

#3. 
##[1]Total cost for meeting the target in county x
##[2]The cost or benifit to each sector
##[3]The tax revenue generated 

##a. Cap on carbon: 300 tons in Country X:Each sector (A, B, and C) must cut its carbon emissions by 100 tons (thus reducing total emissions from 600 down to 300).

```{r}
#3a. 
#new carbon emissions in sector (A = 80, B = 100, C= 120)
mc_a <- sector_a_function(100)
a_x_int <- sector_a_function(0)

tc_a <- (0.5*mc_a*(100-a_x_int))
tc_a
#2663.95

mc_b <- sector_b_function(100)
b_x_int <- sector_b_function(0)

tc_b <- (0.5*mc_b*(100-b_x_int))
tc_b
#1323.61

mc_c <- sector_c_function(100)
c_x_int <- sector_c_function(0)

tc_c <- (0.5*mc_c*(100-c_x_int))
tc_c
#3725.238

tc_total <- sum(tc_a, tc_b, tc_c)
tc_total
#7712.79

# tax revenue is 0 (because there is no tax) 
```

##B. Tax
#Tax on carbon. To emit a ton of carbon in country X, you must pay a tax of $t. You will need to find the tax that accomplishes the desired reduction.

Set a tax where the marginal abatement costs are equal (equamarginal principle)

```{r}
#aggregate demand curve
#switch a slope and int to be in terms of q=

#Sector a slope and intercept (in terms of q) 
a_demand_int1 <- (-y_int_a/a_dem_slope)
a_demand_slope1 <- (1/a_dem_slope)
a_demand1 <- function(x){a_demand_int1 + a_demand_slope1*x}

#Sector b slope and intercept (in terms of q) 
b_demand_int1 <- (-y_int_b/b_dem_slope)
b_demand_slope1 <- (1/b_dem_slope)
b_demand1 <- function(x){b_demand_int1 + b_demand_slope1*x}

#Sector c slope and intercept (in terms of q) 
c_demand_int1 <- (-y_int_c/c_dem_slope)
c_demand_slope1 <- (1/c_dem_slope)
c_demand1 <- function(x){c_demand_int1 + c_demand_slope1*x}

#Sector d slope and intercept (in terms of q) 
d_demand_int1 <- (-y_int_d/d_dem_slope)
d_demand_slope1 <- (1/d_dem_slope)
d_demand1 <- function(x){d_demand_int1 + d_demand_slope1*x}

#Aggregate slope and intercept (in terms of q) (ONLY sectos A, B, C)
agg_demand_int <- sum(a_demand_int1,b_demand_int1,c_demand_int1)
agg_demand_slope <- sum(a_demand_slope1, b_demand_slope1, c_demand_slope1)

#Transformed back in terms of p=
agg_demand_int1 <- (-agg_demand_int/agg_demand_slope)
agg_demand_slope1 <- (1/agg_demand_slope)
agg_demand_curve <- function(x){agg_demand_int1 + agg_demand_slope1*x}

#Determine the optimal tax at 300 tons 
optimal_tax <- agg_demand_curve(300)
optimal_tax
#35.93
```

```{r}
##Plot to determine areas under curves

agg_demand_plot <- ggplot(data.frame(x=c(0,200))) + 
  stat_function(fun=agg_demand_curve, geom="line", aes(color = "Aggregate Demand"))+
  scale_x_continuous(limits =c(0,350), expand = c(0,0))+
  scale_y_continuous(limits=c(0,200), expand=c(0,0))+
  theme_classic() +
  labs(fill = "", 
       color = "Sector",
       x = "Carbon Emissions (tons)",
       y = "Price ($/ton)")+
  ggtitle("Aggregate Demand for Sector A, B, C & D Carbon Emissions")+
  theme_classic() + scale_colour_manual("Sector", values = c("chartreuse4"))

agg_demand_plot
```

```{r}
#emissions from sector A
a_emissions_tax <- a_demand1(optimal_tax)
#102.66 

#cost of reduction for sector A
a_mc_abate <- sector_a_function(180-a_emissions_tax)
a_cost_tax <- (0.5*(180-a_x_int)*a_mc_abate)
a_cost_tax
#2881.77

#emissions from sector B
b_emissions_tax <- b_demand1(optimal_tax)
b_emissions_tax
#48.18

#cost of reduction for sector B
b_mc_abate <- sector_b_function(200-b_emissions_tax)
b_cost_tax <- (0.5*(200-b_x_int)*b_mc_abate)
b_cost_tax
#3765.16

#emissions from sector C
c_emissions_tax <- c_demand1(optimal_tax)
c_emissions_tax
#139.78

#cost of reduction for sector C
c_mc_abate <- sector_c_function(220-c_emissions_tax)
c_cost_tax <- (0.5*(220-c_x_int)*c_mc_abate)
c_cost_tax
#5933.77

total_emissions <- sum(a_emissions_tax, b_emissions_tax, c_emissions_tax)
total_emissions
#300 (GREAT, can tell that is is optimal because it adds up to 300)

cost_total<-sum(a_cost_tax,b_cost_tax,c_cost_tax)
cost_total
#10185.74 (total cost of abatement that meets the equamarginal principle)

tax_rev <- optimal_tax*300
tax_rev
#10780.07 (at optimal tax for sectors A, B, C)
```
"Equimarginal principle: firms will trade until the cost of abatement is the same across firms. We aggregate the demand curves hoirzontally since it is a private good. Since the goal is to abate 300 units of emissions the total cost of abatement that meets the equimarginal principle is 10185.74"

##C.Cap and Trade Country X: Each sector (A, B, C) is allocated carbon permits equal to their current emissions minus 100 (same as in (a) above), thus achieving the total reduction of 300. Then, these three sectors are allowed to trade with each other. You will need to derive the outcome of that trading.

```{r}

#price of optimal tax is equal to price of cap and trade permits
#

a_benefit_trade <- (100-a_emissions_tax)*optimal_tax
a_benefit_trade
#-432.9696 
#benefit emit less than 100 and sell extra permits

b_benefit_trade <- (100-b_emissions_tax)*optimal_tax
b_benefit_trade
#1862.14
#cost because c emits more than 100)

c_benefit_trade <- (100-c_emissions_tax)*optimal_tax
c_benefit_trade
#-1429.17
#benefit emit less than 100 and sell extra permits (most benefits to sector B - lowest MC of abatement and can benefit the most out of all secotrs from abating and trading credits) 

```

#4. Again, without any co-benefits, suppose that country Y (which only has one carbon-emitting sector, D) has no obligation to reduce its emissions. Country X asks country Y to enter the country X carbon market. Doing so would require country Y to put a cap on carbon emissions at its current level of emissions (300 tons), but would then allow country Y to sell carbon offsets to sectors A, B, or C. Are there any incentives for country Y to enter country X’s carbon market and to thus place a voluntary cap on its emissions? Are there any incentives for country X to try to attract country Y into its market?

```{r}
#more about sector d
#Sector d slope and intercept (in terms of q) 
d_demand_int1 <- (-y_int_d/d_dem_slope)
d_demand_slope1 <- (1/d_dem_slope)
d_demand1 <- function(x){d_demand_int1 + d_demand_slope1*x}

#emissions from sector d
d_emissions_tax <- sector_d_function(optimal_tax)
d_emissions_tax
#19.03

#cost of reduction for sector d
d_mc_abate <- sector_d_function(300-d_emissions_tax)
d_mc_abate
#82.72

```


#it would be a benefit to country Y to enter the carbon market because it has a lower cost of abatement of carbon emissions relative to sectors A and C, and WTP for the first units, and therefore would be able to sell/trade carbon credits to Sectors A and C at a benefit to country Y. 

#conversly, country x will want to invite country Y as it would be able to buy permits at a lower cost to emit more carbon without surpassing MC of abatement at emission levels past 100 tons within the market 

#########
#5. Now assume that every ton of carbon emissions creates 1 ton of local air pollution. Local air pollution causes economic damages (health, environmental, etc.), but only in the country in which it is emitted. Assume there are no local air pollution regulations in either country X or country Y.

#a. In a carbon cap and trade market that only covers sectors in country X, how much local air pollution would you expect in country X? In country Y?

#b. If country Y enters the carbon market for country X (as in question 4 above), how much local pollution will there be in country X and country Y?


